---
title: "Import light chest"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

# Preface

This is a work-in-progress descriptive analysis of the `AertsEtAl2025` dataset.

```{r}
#| label: setup
#| include: false
library(LightLogR)
library(glue)
library(tidyverse)
library(gt)
library(readxl)
library(cowplot)
library(legendry)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(patchwork)
library(rlang)
library(here)

remote <- 
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/"

c(
  "filefinder",
  "general_parameters",
  "overview_plot"
) |> walk(\(x) source(paste0(remote, x, ".R")))

site <- "THUAS"
```

# Overview

## Data import: wearable data

The first step is the import of wearable data from the `chest` position (mounted on chest).

```{r}
#| label: "general information"
#regex to extract participant Id and wearing position
# pattern <- "[A-Z]+_S[0-9]{3}_[hcw]"
#regex to extract participant Id
pattern <- "[A-Z]+_S[0-9]{3}"
```

```{r}
files <- filefinder("actlumus_chest", continuous = TRUE, negate = "Report")
```

```{r}
#participant S015 (chest) had a non-synchronized device and needs to be removed from analysis until re-adjusted
data <- import$ActLumus(files[c(1:9,11:15)], tz = tzs[site], auto.id = pattern, dst_adjustment = TRUE)
```

One file need further adjustment, as it contains timestamps from yr 2000. This is a typical issue with `ActLumus` devices after their battery is empty. We will use the other two wearing positions (wrist and head) to derive a valid offset

```{r}
#| fig-height: 20
#| fig-width: 10
#getting all wearing positions
files <- list.dirs("../data/raw/individual/THUAS_S015/continuous", full.names = TRUE)
files <- files[str_detect(files, "actlumus")] |> list.files(full.names = TRUE)
files <- files[str_detect(files, "Report", negate = TRUE)]

#loading the data
pattern <- "[A-Z]+_S[0-9]{3}_[hcw]"
data2 <- 
  import$ActLumus(files, tzs[site], auto.id = pattern,
                  dst_adjustment = TRUE, not.before = "1970-01-01")

data2 |> 
  mutate(
    Datetime = case_when(
      Id == "THUAS_S015_c" ~ Datetime + dyears(25) + dmonths(6) + ddays(5) + dhours(12) + dminutes(32),
      .default = Datetime)
    ) |> 
  gg_day(geom = "line", aes_col = Id, linewidth = 0.5) |> 
  gg_photoperiod(coordinates[[site]])
  # coord_cartesian(xlim = c(22, 24)*3600)

data2 <- 
data2 |> 
  mutate(
    Datetime = case_when(
      Id == "THUAS_S015_c" ~ Datetime + dyears(25) + dmonths(6) + ddays(5) + dhours(12) + dminutes(32),
      .default = Datetime)
    )

```


```{r}
data2 <- 
data2 |> 
  sample_groups(sample = 1) |> 
  mutate(Id = "THUAS_S015")

data <-
  join_datasets(data, data2)
rm(data2)
```

## Regularizing data

In the first step, we will trim the data by the study time.

```{r}
path_study_dates <- paste0("../data/Study_dates_MeLiDos_", site, ".xlsx")

#import table with study times
Study_dates <- read_excel(path_study_dates)
#gather the important information
Study_dates <-
  Study_dates |> 
    rename(Id = subjectID_device, start = datetime_trial_start, end = datetime_trial_end) |> 
    select(Id, start, end) |> 
    mutate(across(c(start, end), \(x) force_tz(x, tzs[site])),
           trial = TRUE) |> 
    filter(str_detect(Id, "_c$")) |>
    mutate(Id = str_remove(Id, "_c$")) |> 
  group_by(Id)

#add the trim information to the dataset and filter by it
data <- 
  data |> 
  add_states(Study_dates) |> 
  dplyr::filter(trial) |> 
  select(-trial)

data |> gg_overview()
```


```{r}
data |> has_gaps()
data |> has_irregulars()
data |> gg_gaps(group.by.days = TRUE, show.irregulars = TRUE, full.days = FALSE)
```

Only one day remains with gap/irregular data. This is due to the daylight savings time jump.

```{r}
#| fig-height: 15
#| fig-width: 5
data_cleaned <- 
data |> 
  gap_handler(full.days = TRUE) #remove implicit gaps

data_cleaned |> has_gaps() #no more implicit gaps
data_cleaned |> has_irregulars() #no more irregular data
data_cleaned |> gap_table(MEDI) |> cols_hide(ends_with("_n"))
```

## Exporting values

```{r}
data_cleaned <- 
  data_cleaned |> mutate(position = "chest")

light_chest_1min <- 
data_cleaned |> 
  aggregate_Datetime("1 minute", numeric.handler = \(x) mean(x, na.rm = TRUE)) |> 
  remove_partial_data(MEDI, threshold.missing = "3 hours", by.date = TRUE)

light_chest <- data_cleaned

save(light_chest_1min, file = "../data/imported/light/light_chest_1minute.RData")
save(light_chest, file = "../data/imported/light/light_chest.RData")
```

## Visualization

```{r}
#| warning: false
#| message: false
#| fig-height: 8
#| fig-width: 12

prefix <- paste0(site, "_")

data_cleaned |>
  mutate(Id = Id |> fct_relabel(\(x) str_remove(x, prefix))) |> 
grand_overview(coordinates[[site]], cities[[site]], countries[[site]], 
               country_colors[[site]], photoperiod_sequence = 1.5)

ggsave("../output/figures/Figure_1.png", width = 17, height = 10, scale = 2, units = "cm")
ggsave("../output/figures/Figure_1.jpeg", width = 17, height = 10, scale = 2, units = "cm")
```

## Stats

### Summary table

```{r}
#| message: false
#| warning: false

table_summary <-
summary_table(
  data_cleaned, 
  coordinates = coordinates[[site]], 
  location = cities[[site]], 
  site = countries[[site]], 
  color = country_colors[[site]],
  histograms = TRUE
)

table_summary

gtsave(table_summary, here("output/tables/table_summary.png"), vwidth = 850, expand = 30)
gtsave(table_summary, here("output/tables/table_summary.pdf"))
gtsave(table_summary |> cols_hide(c(plot)), here("output/tables/table_summary.docx"))
```

